Cubical Types with Positive and Negative Polarities

### 2014-05-19 ring.tex
separate syntax from denotation

For cubical types, their denotations, usually written as $\cubt$, are
$n$-dimensional cubical sets with polarities. An $n$-dimensional cubical set
$\nodet{\cubt_1}{\cubt_2}$ consists of a positive subspace $\cubt_1$ and a
negative subspace $\cubt_2$ (shaded in gray). Each of the subspaces are
cubical set of a lower dimension; the 1-dimensional case reduces to
$\nodet{S}{~}$ which is essentially the denotation of conventional
first-order types. As we explain below, cubical sets are closed under sums,
differences, products.

Examples of 1-dimensional cubical sets include all the sets of the form
$\nodet{S}{~}$ where $S$ is finite set denoting a regular first-order
types $t$. A 2-dimensional cubical set looks like $\nodet{S_1}{S_2}$ which
intuitively corresponds to the difference $t_1 - t_2$ of the two types whose
denotations are $S_1$ and $S_2$ respectively. A 3-dimensional cubical type
looks like $\nodet{(\nodet{S_1}{S_2})}{(\nodet{S_3}{S_4})}$ which intuitively
corresponds to the iterated difference of the appropriate types
$(t_1-t_2)-(t_3-t_4)$ where the successive colors from the root encode the
sign as follows: white in the outermost box corresponds to a positive
polarity and gray in the outermost box corresponds to a negative polarity;
then, as we move inwards the polarities alternate for every change of color.

We now give the denotation of cubical types:
\[\begin{array}{rcl}
\denc{t} &=& \nodet{\den{t}}{~} = 
  \nodet{\nodet{\den{t}}{~}}{~} = \ldots \\
\\
\denc{\tau_1 + \tau_2} &=& \denc{\tau_1} \oplus \denc{\tau_2} \\
\denc{\tau_1 * \tau_2} &=& \denc{\tau_1} \otimes \denc{\tau_2} \\
\denc{- \tau} &=& \ominus \denc{\tau}
\end{array}\]
where the operations $\oplus$, $\otimes$, and $\ominus$ are defined below.
It is possible for cubical types not be balanced in general but it is always
possible to embed an $n$-dimensional set into a larger dimension using the
equivalence in the first clause which formalizes that $t$ is equivalent to
$t-0$ and to $(t-0)-0$ etc.  In the recursive definitions below, we assume
that the types have been extended to be balanced as needed:
\[\begin{array}{rcl}
\ominus~S &=& S \\
\ominus~\nodet{\cubt_1}{\cubt_2} &=& \nodet{\ominus~\cubt_2}{\ominus~\cubt_1} \\
\\
S_1 \oplus S_2 &=& S_1 \uplus S_2 \\
(\nodet{\cubt_1}{\cubt_2}) \oplus (\nodet{\cubt_3}{\cubt_4}) &=& 
  \nodet{\cubt_1 \oplus \cubt_3}{\cubt_2 \oplus \cubt_4} \\
\\
S_1 \otimes S_2 &=& S_1 \times S_2 \\
S \otimes (\nodet{\cubt_1}{\cubt_2}) &=& 
  \nodet{S \otimes \cubt_1}{S \otimes \cubt_2} \\
(\nodet{\cubt_1}{\cubt_2}) \otimes \cubt &=& 
  \nodet{\cubt_1 \otimes \cubt}{\cubt_2 \otimes \cubt}
\end{array}\]

>>>>

For cubical types, their denotations, usually written as $\cubt$, are
$n$-dimensional cubical sets with polarities. An $n$-dimensional cubical set
$\nodet{\cubt_1}{\cubt_2}$ consists of a positive subspace $\cubt_1$ and a
negative subspace $\cubt_2$ (shaded in gray). Each of the subspaces are
cubical set of a lower dimension; the 1-dimensional case reduces to
$\nodet{S}{\phantom{\emptyset}}$ which is essentially the denotation of
conventional first-order types. As we explain below, cubical sets are closed
under sums, differences, products.

Examples of 1-dimensional cubical sets include all the sets of the form
$\nodet{S}{\phantom{\emptyset}}$ where $S$ is finite set denoting a regular
first-order types $t$. A 2-dimensional cubical set looks like
$\nodet{S_1}{S_2}$ which intuitively corresponds to the difference $t_1 -
t_2$ of the two types whose denotations are $S_1$ and $S_2$ respectively. A
3-dimensional cubical type looks like
$\nodet{(\nodet{S_1}{S_2})}{(\nodet{S_3}{S_4})}$ which intuitively
corresponds to the iterated difference of the appropriate types
$(t_1-t_2)-(t_3-t_4)$ where the successive colors from the root encode the
sign as follows: white in the outermost box corresponds to a positive
polarity and gray in the outermost box corresponds to a negative polarity;
then, as we move inwards the polarities alternate for every change of color.

We now give the denotation of cubical types:
\[\begin{array}{rcl}
\denc{t} &=& \nodet{\den{t}}{\phantom{\emptyset}} = 
  \nodet{\nodet{\den{t}}{\phantom{\emptyset}}}{\phantom{\emptyset}} = \ldots \\
\\
\denc{\tau_1 + \tau_2} &=& \denc{\tau_1} \oplus \denc{\tau_2} \\
\denc{\tau_1 * \tau_2} &=& \denc{\tau_1} \otimes \denc{\tau_2} \\
\denc{- \tau} &=& \ominus \denc{\tau}
\end{array}\]
where the operations $\oplus$, $\otimes$, and $\ominus$ are defined below.
It is possible for cubical types not be balanced in general but it is always
possible to embed an $n$-dimensional set into a larger dimension using the
equivalence in the first clause which formalizes that $t$ is equivalent to
$t-0$ and to $(t-0)-0$ etc.  In the recursive definitions below, we assume
that the types have been extended to be balanced as needed:
\[\begin{array}{rcl}
\ominus~S &=& S \\
\ominus~\nodet{\cubt_1}{\cubt_2} &=& \nodet{\ominus~\cubt_2}{\ominus~\cubt_1} \\
\\
S_1 \oplus S_2 &=& S_1 \uplus S_2 \\
(\nodet{\cubt_1}{\cubt_2}) \oplus (\nodet{\cubt_3}{\cubt_4}) &=& 
  \nodet{\cubt_1 \oplus \cubt_3}{\cubt_2 \oplus \cubt_4} \\
\\
S_1 \otimes S_2 &=& S_1 \times S_2 \\
S \otimes (\nodet{\cubt_1}{\cubt_2}) &=& 
  \nodet{S \otimes \cubt_1}{S \otimes \cubt_2} \\
(\nodet{\cubt_1}{\cubt_2}) \otimes \cubt &=& 
  \nodet{\cubt_1 \otimes \cubt}{\cubt_2 \otimes \cubt}
\end{array}\]
